Nice people or potential cooperators when keeping promises

### 2019-08-09 behavioral-promises/behavioral-promises.tex
Using springer template

\documentclass[12pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Nice people or potential cooperators when keeping promises: an experimental and Bayesian account for two explanations},
            pdfauthor={Said Jiménez},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\providecommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}

  \title{Nice people or potential cooperators when keeping promises: an
experimental and Bayesian account for two explanations}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
    \author{Said Jiménez}
    \preauthor{\centering\large\emph}
  \postauthor{\par}
      \predate{\centering\large\emph}
  \postdate{\par}
    \date{August 2019}

@@

\begin{document}
\maketitle

@@

Human are social animals and we get survival advantages from being such.
Historically, belonging to a social group has provided protection and
guaranteed access to resources such as food, and currently, there has
been evidence that belonging to a social group has a positive effect on
longevity, physical and mental health (Holt-Lunstad, 2018; Tough,
Siegrist, \& Fekete, 2017). The relationship between social closeness
and cooperation is key to the formation of large groups of individuals
without a genetic relationship, and such groups form the basis of
communities, societies, and nations, arguably constituting one of the
most fundamental conditions for human survival (Fehr \& Fischbacher,
2003; Fehr \& Schurtenberger, 2018). An important requirement for
cooperation between individuals is communication. A meta-analysis of 45
studies reported a large positive effect (\(d = 1.01\)) of communication
in cooperation regardless of the communication medium (Balliet, 2010).
In parallel, it is known that after communication sessions between
completely unknown individuals there is an increase in subjective social
closeness (Aron, Melinat, Aron, Vallone, \& Bator, 1997). Thus, these
findings suggest a positive relationship between communication,
cooperation, and social closeness.

One of the forms of communication that have received the most attention
in psychological game theory is ``the promise''. In an interaction
between sender and recipient, it is believed that promises influence
beliefs of the recipient, generating trust and cooperation (Charness \&
Dufwenberg, 2006; Vanberg, 2008). However, there are situations where
trust is betrayed, promises are broken or people deceived. The General
Social Survey (GSS) in 2018 showed that approximately 11\% of people
surveyed responded that they have had sex with someone different to
their partner while they are married (Smith, Davern, Freese, \& Hout,
2018). In economic and social terms, the United Nations Organization has
stressed that deception (corruption) has a global cost of at least 5
percent of the annual gross domestic product (GDP), and it is a
generator of conflict and instability in all nations (UN, 2018).

Despite the widespread presence of interactions between human belonging
to the same group, it is not known how social closeness between
participants can affect keeping or breaking promises. This is important
because it has been pointed out that lies and betrayals of trust occur
daily in close relationships, such as friends, workmates, or classmates.
And, their consequences can be serious such as loss of the relationship,
work or damage to reputation (DePaulo, Ansfield, Kirkendol, \& Boden,
2004). In the laboratory, these transgressions to trust have been
explicitly studied in experiments where subjects have incentives to lie
or not keep their promises. However, the vast majority of these studies
have been conducted in people who do NOT know each other (Baumgartner,
Fischbacher, Feierabend, Lutz, \& Fehr, 2009; Baumgartner, Gianotti, \&
Knoch, 2013; Charness \& Dufwenberg, 2006; Fischbacher \& Föllmi-Heusi,
2013; Gneezy, 2005; Gneezy, Rockenbach, \& Serra-Garcia, 2013; Mazar,
Amir, \& Ariely, 2008). Some studies address deception in close
interpersonal relationships using self-report measurements, however,
they have limitations due to social desirability bias (DePaulo et al.,
2004; DePaulo \& Kashy, 1998) or rather explore the development of
deception detection skills in same-sex friends (Anderson, DePaulo, \&
Ansfield, 2002).

In this study, we wanted to understand the effects of three partners
with different levels of social closeness on keeping or breaking
promises, as well as on cooperation, using an experimental trust game.
We analyzed it in light of the two main motivations that have been
pointed out in the literature regarding keeping promises which are:
instrumental and intrinsic (Baumgartner et al., 2009):

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Social closeness will reduce the decision of breaking the promise.
  According to the \emph{instrumental} motivation, we expect that
  subjects keep their promises with friends for the purpose of
  facilitating future cooperation, which can be extended even beyond the
  trials in the experiment. We also expect, although to a lesser extent,
  that the subjects keep promises to strangers, with the purpose of
  facilitating cooperation at least during the trials during the
  experiment. Finally, we anticipate that the participants break the
  promises to the computer because it is a partner without social
  closeness and they could not ensure cooperation in future trials. It
  should be noted that if the participants keep the promises to the
  computer, evidence would be given in favor of \emph{intrinsic}
  motivation.
\item
  There will be an effect of social closeness on cooperation, regardless
  of promises. We expect that there is more probability of paying the
  friend than the stranger and more probability of paying the stranger
  than the computer. According to the \emph{instrumental motivation},
  the cooperation will be greater for partners with whom the subject
  anticipates greater cooperation in the future (friend \textgreater{}
  stranger \textgreater{} computer).
\item
  Finally, we will obtain two subsamples according to the cooperation
  rate of the subjects as was done in another study (Baumgartner et al.,
  2009). However, what in the aforementioned study was classified as a
  group of ``honest'' and ``dishonest'', we will show that in our sample
  it is not supported because if the groups differ in their payment rate
  it will be due to the difference in their level of commitment
  expressed by promises.
\end{enumerate}

@@

All subjects performed an adaptation of the trust game with promises,
using hypothetical monetary rewards (Baumgartner et al., 2009). The
variation with respect to the original task is that in our experiment
three partners were presented with three levels of social closeness in
the role of the investor: computer (no closeness), strange (low
closeness) and friend (high closeness), while the participant acted as
the trustee. The task was programmed in PsychoPy2 version 1.84.2 (Peirce
et al., 2019; Peirce, 2008) and consisted of 24 trials between two
players: trustee and investor. The trustee originally has 0 Mexican
pesos and the investor has 2 pesos, the investor is presented with the
opportunity to give his money to the trustee or keep it. If investor
gives his money, it is multiplied by 5 pesos, so that the trustee has 10
pesos. Finally, the trustee decides to pay half to the investor or keep
the 10 pesos. The structure described is repeated in 24 trials, however,
in 4 of the trials the trustee can send a promise to the investor, the
promises were that always, mostly, sometimes, or never would pay back.
Each promise was valid for three trials so that 12 of the trials have
the effect of the promise and the other 12 do not. Each partner performs
the role of investor in 8 of 24 total trials, however, both the promises
and social closeness conditions were presented to the trustee in
pseudorandom order.

@@

Subjects came to the lab with a same-sex friend considered close by him
or her. It was emphasized that they must not have a romantic or family
relationship with their friends, to try to exclude the effect on
cooperation due to a consanguineous relationship or sexual attraction.
Also, to ensure that the subjects and their companions had a similar
degree of social closeness to each other, both responded the ``Inclusion
of the Other in Self'' IOS scale (Aron et al., 1997) without observing
their partner responses. The scale consists of seven pairs of circles
that vary in the degree of overlap between them, the respondent must
select the pair of circles that best represents the subjective closeness
to his partner. Subjects were trained in the main task by two
researchers, the game was first explained verbally, and then the
computer interface was presented on a portable computer. They were told
that they would make their decisions with four buttons on the computer
keyboard, which corresponded to the levels of promises and with only two
of these to make their decision to pay back or not. Friends heard the
explanation, they were told that they would make their decisions in a
separate room on another computer even though they were not. Subjects
performed between 3 and 6 practice tests together with the researcher
who exemplified the course of the game when the investor gave his
budget. When the subjects were ready, we proceeded to take them without
their companions to another room (or MRI scanner it that was the case)
where they would perform the task.The subjects began the task believing
that they would really play with the three partners of different social
closeness. In the other room, we debriefed the companion about the
deception. When the subject finished the task, they were debriefed.
Although no participant exercised their right, both subjects and their
friends were told that they could withdraw their participation and
informed consent if they considered disagreeing with any of the
manipulations made by the researchers.

@@

All models presented below were programmed in R via the \textbf{brms}
package (Bürkner, 2017; R Core Team, 2019), which performs the inference
using sampling by Markov chain Monte Carlo through \textbf{Stan} (Stan
Development Team, 2018). For each model, the posteriors distributions of
all parameters were approximated with four chains of 2000 iterations
each, the first 1000 iterations of each chain were discarded (burning
period), for a total of 4000 post burning samples. Models convergence
was evaluated through visual inspection of the chains and calculation of
the \(\hat{R}\) statistic, which for all parameters was 1, that can be
interpreted as convergence. Vaguely informative prior distributions were
used for the parameters of interest in the models, which allows the data
to dominate the inference, also assumes as little as possible regarding
the nature of the phenomenon, which could be adequate for the current
state of evidence in the problem we are studying (McElreath, 2018).

@@

{\centering \includegraphics[width=0.8\linewidth]{article_english_files/figure-latex/fig2-1}

@@

To assess social closeness, 30 (15 men) of our subjects and their
friends responded to the IOS scale (Aron et al., 1997; Fareri \&
Delgado, 2014; Sip, Smith, Porcelli, Kar, \& Delgado, 2015), which
consists of seven pairs of circles that vary in the degree of overlap
and represent the social closeness that an individual perceives with
respect to the other. A Hierarchical and Cumulative Bayesian Model (See
section: Commitment expressed in the promises), supports the hypothesis
that there are no differences in subjective social closeness between
subjects and their friends, with a posterior evidence ratio of 2.74 in
its favor.

@@

To evaluate the hypothesis that social closeness would reduce the
behavior of breaking the promise, we filtered all trials in which
subjects promised that they would always pay back and then decided not
to pay. Subsequently, Bayesian inference was used to assess the effect
of partners on the decision to break the promise at the individual and
population level. For this purpose, a Hierarchical Bayesian Model was
carried out that assumes that the uncertainty in partners effect on the
decision to break the promise varies depending on each individual,
however, it also assumes that these variations belong to common
population distributions (Gelman \& Hill, 2006).

@@

In the previous model, the decision to break the promise \(y_i\) comes
from a Bernoulli distribution with probability \(\theta_i\), the goal of
the Hierarchical Model is to predict each decision through the linear
combination of the effects of each partner, transformed by its inverse
link function \(\mathrm{logit}\) (Bürkner, 2017). In this model,
\(\beta\) and \(u\) are coefficients at the population level and
individual level respectively, while \(\mathbf{X}\), \(\mathbf{Z}\) are
their corresponding design matrices. In this case, the population
coefficients correspond to the presence of partners with no social
closeness \(\beta^{computer}\), low \(\beta^{stranger}\) and high
\(\beta^{friend}\).

@@

{\centering \includegraphics[width=0.8\linewidth]{article_english_files/figure-latex/fig3-1}

@@

{\centering \includegraphics[width=0.8\linewidth]{article_english_files/figure-latex/fig4-1}

@@

Figure 5 shows the posterior predictive distribution in contrast to the
proportion of decisions to each partner during the promises phase, and
each panel corresponds to one of the first twelve subjects. The
responses on the dotted line would indicate that the subject paid back
random to that partner. Posterior predictive distribution simulates
observations of the model and compares them with the actual data, it
helps us to identify if the model is sufficiently close to the process
that generated the data (Lee \& Wagenmakers, 2014; Schad, Betancourt, \&
Vasishth, 2019). It can be seen that there is a great correspondence
between the subject's responses and predictions of the model. Even in
cases where the model is ``wrong'' (for example, subject 1), the
observed response is in the range of a predicted standard deviation,
which gives credibility to estimates.

@@

{\centering \includegraphics[width=1\linewidth,height=1\textheight]{article_english_files/figure-latex/fig5-1}

@@

{\centering \includegraphics[width=0.8\linewidth]{article_english_files/figure-latex/fig6-1}

@@

\begin{table}[!h]

\caption{\label{tab:tabla}Posterior coefficients estimates}
\centering
\begin{tabular}{lrrrr}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{2}{c}{ } & \multicolumn{2}{c}{95 \% CI} \\
\cmidrule(l{3pt}r{3pt}){4-5}
Term & Estimate & Est.Error & Lower & Upper\\
\midrule
Intercept & -0.153 & 0.536 & -1.005 & 0.617\\
Stranger & 0.829 & 0.410 & 0.128 & 1.470\\
Friend & 2.084 & 1.248 & -0.151 & 3.975\\
\bottomrule
\end{tabular}
\end{table}

@@

In the original study of promises, authors chose to divide their sample
into two according to the hierarchical clustering technique with Ward's
method (Baumgartner et al., 2009). In this way they obtained two sets of
participants that were different in their payment rates, despite the
fact that they both made very high promises. For this reason, the
authors named the group that paid little as ``dishonest'' and the group
that paid a lot as ``honest''. In a similar exercise, in the present
study we performed the hierarchical grouping technique to obtain a
solution for two groups and we found two similar sets in n that we call
the group ``Low'' and ``High'' (Low = 25, High = 20), with quite
different payment ratios (Low = 58\%, High = 83\%), and a 95\%
confidence interval of 20\% to 32\% in the difference favor of the High
group.

>>>>

\AtBeginDocument{%
  \paperwidth=\dimexpr
    1in + \oddsidemargin
    + \textwidth
    % + \marginparsep + \marginparwidth
    + 1in + \oddsidemargin
  \relax
  \paperheight=\dimexpr
    1in + \topmargin
    + \headheight + \headsep
    + \textheight
    % + \footskip
    + 1in + \topmargin
  \relax
  \usepackage[pass]{geometry}\relax
}

% !TeX program = pdfLaTeX
\documentclass[smallextended]{svjour3}       % onecolumn (second format)
%\documentclass[twocolumn]{svjour3}          % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof
%
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[utf8]{inputenc}

\usepackage[hyphens]{url} % not crucial - just used below for the URL
\usepackage{hyperref}
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

%
% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
%\usepackage{latexsym}
% etc.
%
% please place your own definitions here and don't use \def but
% \newcommand{}{}
%
% Insert the name of "your journal" with
% \journalname{myjournal}
%

%% load any required packages here

@@

\begin{document}

\title{Nice people or potential cooperators when keeping promises }
 \subtitle{An experimental and Bayesian account for two explanations}

    \titlerunning{Nice people or potential cooperators}

\author{  Said Jiménez \and  Kamila E. Sip \and  Roberto E. Mercadillo \and  Diego Angeles-Valdez \and  Jairo Muñoz-Delgado \and  Juan J. Sánchez-Sosa \and  Eduardo A. Garza-Villarreal \and  }

    \authorrunning{ S. Jiménez, et al. }

\institute{
        Said Jiménez \at
     Department of Psychology, National University of México (UNAM) \\
     \email{\href{mailto:said.ejp@gmail.com}{\nolinkurl{said.ejp@gmail.com}}}  %  \\
%             \emph{Present address:} of F. Author  %  if needed
    \and
        Kamila E. Sip \at
     Institute for Neuroleadership \\
     \email{\href{mailto:kami.sip@gmail.com}{\nolinkurl{kami.sip@gmail.com}}}  %  \\
%             \emph{Present address:} of F. Author  %  if needed
    \and
        Roberto E. Mercadillo \at
     Department of Biology, Universidad Autónoma Metropolitana (UAM) \\
     \email{\href{mailto:emmanuele.mercadillo@gmail.com}{\nolinkurl{emmanuele.mercadillo@gmail.com}}}  %  \\
%             \emph{Present address:} of F. Author  %  if needed
    \and
        Diego Angeles-Valdez \at
     Institute of Neurobiology, UNAM \\
     \email{\href{mailto:anglsd_20@hotmail.com}{\nolinkurl{anglsd\_20@hotmail.com}}}  %  \\
%             \emph{Present address:} of F. Author  %  if needed
    \and
        Jairo Muñoz-Delgado \at
     Investigaciones en Neurociencias, INPRF \\
     \email{\href{mailto:munozd460@gmail.com}{\nolinkurl{munozd460@gmail.com}}}  %  \\
%             \emph{Present address:} of F. Author  %  if needed
    \and
        Juan J. Sánchez-Sosa \at
     Department of Psychology, National University of México (UNAM) \\
     \email{\href{mailto:jujosaso@gmail.com}{\nolinkurl{jujosaso@gmail.com}}}  %  \\
%             \emph{Present address:} of F. Author  %  if needed
    \and
        Eduardo A. Garza-Villarreal \at
     Institute of Neurobiology, UNAM \\
     \email{\href{mailto:egarza@gmail.com}{\nolinkurl{egarza@gmail.com}}}  %  \\
%             \emph{Present address:} of F. Author  %  if needed
    \and
    }

\date{Received: date / Accepted: date}
% The correct dates will be entered by the editor

\maketitle

\begin{abstract}
Promises increase cooperation between non-genetically related
individuals, however, lies and betrayals occur daily in close social
relationships. Despite this, the effect of social closeness on the
decision to keep or break a promise has not been studied. We conducted
an experiment in which subjects could freely decide if they broke or
kept their promise to three partners with different levels of social
closeness: null (computer), low (strange) and high (friend). If subjects
keep their word with the computer, it provides evidence in favor of
intrinsic motivation (comply to do the correct thing), otherwise
extrinsic motivation would be supported (comply to facilitate future
cooperation). Using Hierarchical Bayesian Modeling, we find that if
social closeness increases, the probability of breaking promises
decreases monotonically, the evidence ratio in favor to the hypothesis
stranger \textless{} computer was 7.47, and friend \textless{} stranger
was 8.64. Furthermore, using Bayesian Cumulative Modeling we find that
subjects were very consistent between their promises and the subsequent
decisions they make. Although is more probable that our subjects keep
their promises to any of the three partners, there is a proportion of
transgressions to the computer, whose inference excludes zero with a
95\% posterior probability. Also, subjects never broke the promise to
their friends. Our results indicate that subjects are predominantly
``nice people'' because they keep their promises even to partners
without social closeness, however, if the subjects are not certain that
their partner will also be a ``potential cooperator'', dishonesty
emerges in the form of broken promises.
\\
\keywords{
        Deception \and
        Hierarchical Bayesian Models \and
        Bayesian Cumulative Models \and
        Social closeness \and
        Promises \and
    }

\end{abstract}

\def\spacingset#1{\renewcommand{\baselinestretch}%
{#1}\small\normalsize} \spacingset{1}

@@

Human are social animals and we get survival advantages from being such.
Historically, belonging to a social group has provided protection and
guaranteed access to resources such as food, and currently, there has
been evidence that belonging to a social group has a positive effect on
longevity, physical and mental health (Tough, Siegrist, and Fekete 2017;
Holt-Lunstad 2018). The relationship between social closeness and
cooperation is key to the formation of large groups of individuals
without a genetic relationship, and such groups form the basis of
communities, societies, and nations, arguably constituting one of the
most fundamental conditions for human survival (Fehr and Fischbacher
2003; Fehr and Schurtenberger 2018). An important requirement for
cooperation between individuals is communication. A meta-analysis of 45
studies reported a large positive effect (\(d = 1.01\)) of communication
in cooperation regardless of the communication medium (Balliet 2010). In
parallel, it is known that after communication sessions between
completely unknown individuals there is an increase in subjective social
closeness (Aron et al. 1997). Thus, these findings suggest a positive
relationship between communication, cooperation, and social closeness.

One of the forms of communication that have received the most attention
in psychological game theory is ``the promise''. In an interaction
between sender and recipient, it is believed that promises influence
beliefs of the recipient, generating trust and cooperation (Charness and
Dufwenberg 2006; Vanberg 2008). However, there are situations where
trust is betrayed, promises are broken or people deceived. The General
Social Survey (GSS) in 2018 showed that approximately 11\% of people
surveyed responded that they have had sex with someone different to
their partner while they are married (Smith et al. 2018). In economic
and social terms, the United Nations Organization has stressed that
deception (corruption) has a global cost of at least 5 percent of the
annual gross domestic product (GDP), and it is a generator of conflict
and instability in all nations (UN 2018).

Despite the widespread presence of interactions between human belonging
to the same group, it is not known how social closeness between
participants can affect keeping or breaking promises. This is important
because it has been pointed out that lies and betrayals of trust occur
daily in close relationships, such as friends, workmates, or classmates.
And, their consequences can be serious such as loss of the relationship,
work or damage to reputation (DePaulo et al. 2004). In the laboratory,
these transgressions to trust have been explicitly studied in
experiments where subjects have incentives to lie or not keep their
promises. However, the vast majority of these studies have been
conducted in people who do NOT know each other (Gneezy 2005; Gneezy,
Rockenbach, and Serra-Garcia 2013; Baumgartner et al. 2009; Baumgartner,
Gianotti, and Knoch 2013; Charness and Dufwenberg 2006; Mazar, Amir, and
Ariely 2008; Fischbacher and Föllmi-Heusi 2013). Some studies address
deception in close interpersonal relationships using self-report
measurements, however, they have limitations due to social desirability
bias (DePaulo and Kashy 1998; DePaulo et al. 2004) or rather explore the
development of deception detection skills in same-sex friends (Anderson,
DePaulo, and Ansfield 2002).

In this study, we wanted to understand the effects of three partners
with different levels of social closeness on keeping or breaking
promises, as well as on cooperation, using an experimental trust game.
We analyzed it in light of the two main motivations that have been
pointed out in the literature regarding keeping promises which are:
instrumental and intrinsic (Baumgartner et al. 2009):

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Social closeness will reduce the decision of breaking the promise.
  According to the \emph{instrumental} motivation, we expect that
  subjects keep their promises with friends for the purpose of
  facilitating future cooperation, which can be extended even beyond the
  trials in the experiment. We also expect, although to a lesser extent,
  that the subjects keep promises to strangers, with the purpose of
  facilitating cooperation at least during the trials during the
  experiment. Finally, we anticipate that the participants break the
  promises to the computer because it is a partner without social
  closeness and they could not ensure cooperation in future trials. It
  should be noted that if the participants keep the promises to the
  computer, evidence would be given in favor of \emph{intrinsic}
  motivation.
\item
  There will be an effect of social closeness on cooperation, regardless
  of promises. We expect that there is more probability of paying the
  friend than the stranger and more probability of paying the stranger
  than the computer. According to the \emph{instrumental motivation},
  the cooperation will be greater for partners with whom the subject
  anticipates greater cooperation in the future (friend \textgreater{}
  stranger \textgreater{} computer).
\item
  Finally, we will obtain two subsamples according to the cooperation
  rate of the subjects as was done in another study (Baumgartner et al.
  2009). However, what in the aforementioned study was classified as a
  group of ``honest'' and ``dishonest'', we will show that in our sample
  it is not supported because if the groups differ in their payment rate
  it will be due to the difference in their level of commitment
  expressed by promises.
\end{enumerate}

@@

All subjects performed an adaptation of the trust game with promises,
using hypothetical monetary rewards (Baumgartner et al. 2009). The
variation with respect to the original task is that in our experiment
three partners were presented with three levels of social closeness in
the role of the investor: computer (no closeness), strange (low
closeness) and friend (high closeness), while the participant acted as
the trustee. The task was programmed in PsychoPy2 version 1.84.2 (Peirce
et al. 2019; Peirce 2008) and consisted of 24 trials between two
players: trustee and investor. The trustee originally has 0 Mexican
pesos and the investor has 2 pesos, the investor is presented with the
opportunity to give his money to the trustee or keep it. If investor
gives his money, it is multiplied by 5 pesos, so that the trustee has 10
pesos. Finally, the trustee decides to pay half to the investor or keep
the 10 pesos. The structure described is repeated in 24 trials, however,
in 4 of the trials the trustee can send a promise to the investor, the
promises were that always, mostly, sometimes, or never would pay back.
Each promise was valid for three trials so that 12 of the trials have
the effect of the promise and the other 12 do not. Each partner performs
the role of investor in 8 of 24 total trials, however, both the promises
and social closeness conditions were presented to the trustee in
pseudorandom order.

@@

Subjects came to the lab with a same-sex friend considered close by him
or her. It was emphasized that they must not have a romantic or family
relationship with their friends, to try to exclude the effect on
cooperation due to a consanguineous relationship or sexual attraction.
Also, to ensure that the subjects and their companions had a similar
degree of social closeness to each other, both responded the ``Inclusion
of the Other in Self'' IOS scale (Aron et al. 1997) without observing
their partner responses. The scale consists of seven pairs of circles
that vary in the degree of overlap between them, the respondent must
select the pair of circles that best represents the subjective closeness
to his partner. Subjects were trained in the main task by two
researchers, the game was first explained verbally, and then the
computer interface was presented on a portable computer. They were told
that they would make their decisions with four buttons on the computer
keyboard, which corresponded to the levels of promises and with only two
of these to make their decision to pay back or not. Friends heard the
explanation, they were told that they would make their decisions in a
separate room on another computer even though they were not. Subjects
performed between 3 and 6 practice tests together with the researcher
who exemplified the course of the game when the investor gave his
budget. When the subjects were ready, we proceeded to take them without
their companions to another room (or MRI scanner it that was the case)
where they would perform the task.The subjects began the task believing
that they would really play with the three partners of different social
closeness. In the other room, we debriefed the companion about the
deception. When the subject finished the task, they were debriefed.
Although no participant exercised their right, both subjects and their
friends were told that they could withdraw their participation and
informed consent if they considered disagreeing with any of the
manipulations made by the researchers.

@@

All models presented below were programmed in R via the \textbf{brms}
package (Bürkner 2017; R Core Team 2019), which performs the inference
using sampling by Markov chain Monte Carlo through \textbf{Stan} (Stan
Development Team 2018). For each model, the posteriors distributions of
all parameters were approximated with four chains of 2000 iterations
each, the first 1000 iterations of each chain were discarded (burning
period), for a total of 4000 post burning samples. Models convergence
was evaluated through visual inspection of the chains and calculation of
the \(\hat{R}\) statistic, which for all parameters was 1, that can be
interpreted as convergence. Vaguely informative prior distributions were
used for the parameters of interest in the models, which allows the data
to dominate the inference, also assumes as little as possible regarding
the nature of the phenomenon, which could be adequate for the current
state of evidence in the problem we are studying (McElreath 2018).

@@

{\centering \includegraphics[width=0.8\linewidth]{behavioral-promises_files/figure-latex/fig2-1}

@@

To assess social closeness, 30 (15 men) of our subjects and their
friends responded to the IOS scale (Aron et al. 1997; Fareri and Delgado
2014; Sip et al. 2015), which consists of seven pairs of circles that
vary in the degree of overlap and represent the social closeness that an
individual perceives with respect to the other. A Hierarchical and
Cumulative Bayesian Model (See section: Commitment expressed in the
promises), supports the hypothesis that there are no differences in
subjective social closeness between subjects and their friends, with a
posterior evidence ratio of 2.74 in its favor.

@@

To evaluate the hypothesis that social closeness would reduce the
behavior of breaking the promise, we filtered all trials in which
subjects promised that they would always pay back and then decided not
to pay. Subsequently, Bayesian inference was used to assess the effect
of partners on the decision to break the promise at the individual and
population level. For this purpose, a Hierarchical Bayesian Model was
carried out that assumes that the uncertainty in partners effect on the
decision to break the promise varies depending on each individual,
however, it also assumes that these variations belong to common
population distributions (Gelman and Hill 2006).

@@

In the previous model, the decision to break the promise \(y_i\) comes
from a Bernoulli distribution with probability \(\theta_i\), the goal of
the Hierarchical Model is to predict each decision through the linear
combination of the effects of each partner, transformed by its inverse
link function \(\mathrm{logit}\) (Bürkner 2017). In this model,
\(\beta\) and \(u\) are coefficients at the population level and
individual level respectively, while \(\mathbf{X}\), \(\mathbf{Z}\) are
their corresponding design matrices. In this case, the population
coefficients correspond to the presence of partners with no social
closeness \(\beta^{computer}\), low \(\beta^{stranger}\) and high
\(\beta^{friend}\).

@@

{\centering \includegraphics[width=0.8\linewidth]{behavioral-promises_files/figure-latex/fig3-1}

@@

{\centering \includegraphics[width=0.8\linewidth]{behavioral-promises_files/figure-latex/fig4-1}

@@

Figure 5 shows the posterior predictive distribution in contrast to the
proportion of decisions to each partner during the promises phase, and
each panel corresponds to one of the first twelve subjects. The
responses on the dotted line would indicate that the subject paid back
random to that partner. Posterior predictive distribution simulates
observations of the model and compares them with the actual data, it
helps us to identify if the model is sufficiently close to the process
that generated the data (Schad, Betancourt, and Vasishth 2019; Lee and
Wagenmakers 2014). It can be seen that there is a great correspondence
between the subject's responses and predictions of the model. Even in
cases where the model is ``wrong'' (for example, subject 1), the
observed response is in the range of a predicted standard deviation,
which gives credibility to estimates.

@@

{\centering \includegraphics[width=1\linewidth,height=0.42\textheight]{behavioral-promises_files/figure-latex/fig5-1}

@@

{\centering \includegraphics[width=0.8\linewidth]{behavioral-promises_files/figure-latex/fig6-1}

@@

\begin{table}[t]

\caption{\label{tab:tabla}Posterior coefficients estimates}
\begin{tabular}{lrrrr}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{2}{c}{ } & \multicolumn{2}{c}{95 \% CI} \\
\cmidrule(l{3pt}r{3pt}){4-5}
Term & Estimate & Est.Error & Lower & Upper\\
\midrule
Intercept & -0.153 & 0.536 & -1.005 & 0.617\\
Stranger & 0.829 & 0.410 & 0.128 & 1.470\\
Friend & 2.084 & 1.248 & -0.151 & 3.975\\
\bottomrule
\end{tabular}
\end{table}

@@

In the original study of promises, authors chose to divide their sample
into two according to the hierarchical clustering technique with Ward's
method (Baumgartner et al. 2009). In this way they obtained two sets of
participants that were different in their payment rates, despite the
fact that they both made very high promises. For this reason, the
authors named the group that paid little as ``dishonest'' and the group
that paid a lot as ``honest''. In a similar exercise, in the present
study we performed the hierarchical grouping technique to obtain a
solution for two groups and we found two similar sets in n that we call
the group ``Low'' and ``High'' (Low = 25, High = 20), with quite
different payment ratios (Low = 58\%, High = 83\%), and a 95\%
confidence interval of 20\% to 32\% in the difference favor of the High
group.

Although it seems a similar result to that reported in that paper
(Baumgartner et al. 2009), we explore the pattern of promises of both
groups to determine if it was possible to classify our subjects as
honest and dishonest. If we hypothesize that the commitment to pay would
be reflected in the level of promises selected, a group of dishonest
people could generate in their partners the belief that they will pay by
choosing a high level of promises (mostly or always) and, later,
betraying that trust when deciding not to pay back. In order to explore
whether the groups represent populations that do not differ in the level
of commitment expressed in the promises, we use a Cumulative Bayesian
Model, which assumes that the levels of promises are an observed ordinal
variable \(Y\) that originates from the categorization of a continuous
latent variable \(\tilde{Y}\), for this case, the expressed commitment
to pay back (Bürkner and Vuorre 2019). The degree to which the subjects
of the High group differ from the Low group, in normal standard
deviations (\(z\)-values), on the latent scale of \(\tilde{Y}\), has a
point estimate of 0.95, which implies that the High group has 0.95
z-values greater commitment to pay back than the Low group. The 95\%
Bayesian Credibility Interval indicates that the High group is between
0.40 to 1.51 \(z\)-values of difference from Low. So we can conclude
with at least a 95\% probability that people belonging to the High group
expressed in their promises a greater commitment to pay back than the
subjects of the Low group. If we look at Figure 7, the probability of
choosing the different levels of promises varies depending on the group,
the one who had the highest percentage of decisions to pay back is more
likely to choose always (High group), while the group that had the
lowest percentage of decisions to pay back are more likely to choose
mostly (Low group). According to our data, we could not justify the
classification of our groups according to their honesty or dishonesty,
at least not with the technique used in Baumgartner et al. (2009). Since
people were quite consistent with keeping what they promised to pay.
